Landau quantization of nearly degenerate bands

### 2018-09-29 paper.tex
Updates from ShareLaTeX

In the absence of magnetic breakdown, semiclassical quantization rules are well-known for Landau levels of  both nondegenerate [Onsager, Lifshitz] and degenerate [Roth, Mikitik, Topoferm, 100page] energy bands. Our work generalizes these rules to nearly-degenerate bands, whose orbits are split in $\bk$-space with a differential area  $\delta S$ that is much less than the average area $S$ of the orbits; we shall refer to $\delta S/S{\ll}1$ as the \textit{quasidegeneracy} condition. Quasidegenerate orbits commonly originate from spin-orbit-splitting of energy bands in noncentrosymmetric solids. We provide an illustrative model of a two dimensional electron gas (2DEG) with Rashba spin-orbit coupling:
\begin{equation}
H_R=\frac{{\hbar^2} k^2}{2m}+\alpha  (k_{x}\sigma_{y}-k_{y}\sigma_{x}).\label{eq:Rashba-Hamiltonian}
\end{equation}
At energies much greater than $m\alpha^2$, the differential area of the two spin-orbit-split orbits [see Fig. \ref{fig:orbits}(a)] is much smaller than their average area:\footnote{The expressions for $\delta S$ and $S$ in \q{deltaSvsS} are asymptotically valid for $E{\gg}m\alpha^2$.}
\e{\delta S = 2\pi k_E \f{2\alpha m}{\hbar^2} \ll S =\pi k_E^2; \;\; k_{E}:=\f{\sqrt{2m E}}{\hbar} .\la{deltaSvsS}}
$S$ is equivalently the area of the circular, constant-energy contour in the absence of spin-orbit coupling ($\alpha{=}0$).

@@

The quantization rule for Landau levels reflects the single-valuedness of the semiclassical (WKB) wavefunction over $\frako_0$,[Berry, Mount] i.e.,  the phase accumulated by an electron (over one cyclotron period $T_c$) is an integer multiple of $2\pi$: $(l^2S(E){+}\gamma)/2\pi{\in}\Z$. The leading-order phase   involves the classical action $S{:}{=}-\int_{0}^{T_c} k_x \dot{k}_y dt$  corresponding to $(K_x,K_y)$;[mount] $S$ has a dual interpretation as the signed area (in $\bk$-space) enclosed by $\frako_0$, as illustrated in Fig. \ref{fig:orbits}(b). $\gamma$ is a subleading Maslov correction [caustic] that generally equals $\pi$ times the rotation number $r$ of $\frako_0$ (i.e., the number of rotations made by the tangent vector over one cycle);[100page] for orbits that are deformable to a circle,  $r{=}{\pm} 1$. In the zeroth-order  solution, all Landau levels are spin-degenerate with a level spacing 
\e{ \var_c:=\f{2\pi}{l^2|\partial S/\partial E|}:=\f{\hbar^2}{ml^2}:=\f{h}{T_c}. \la{def:cyclotron}} 
Here, $\var_c$ and $T_c$ are respectively the cyclotron energy and period of the zeroth-order orbit, and $m$ is the effective mass. Henceforth, `degeneracy of Landau levels' should be understood as a spin degeneracy.

@@

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta \hat{H}$ to the degenerate band subspace of $\hat{H}_0$. In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively), $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,\delta \var_-$, corresponding to outer and inner orbit respectively; $\delta \var_+<\delta \var_-$) are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of  \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \la{spinsplitwkb}}
%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

The remaining terms in \q{eq:H1} describe a generalized Zeeman interaction $B\calm$ which is the first-order-in-$B$ correction to the Peierls-Onsager Hamiltonian $\var(\bK)$.[kohn,blount,roth] Included in this interaction is the Zeeman coupling to the spin ($g_s\mu_Bs_z/\hbar$) and orbital ($M_z$) magnetic moment of the quasidegenerate band; the latter is expressible in terms of matrix elements of the velocity operator $\boldsymbol{\Pi}_{mn}=i\langle\psi_m|[\hat{H}, \hat{\boldsymbol{r}}]|\psi_n\rangle/\hbar$ between Bloch wave functions $\psi_{n\bk}$:
\e{
M_z=\f{ie\hbar}{2}\sum_l'\f{\Pi_{x,ml}\Pi_{y,ln}-\Pi_{y,ml}\Pi_{x,ln}}{\epsilon_m-\epsilon_l}. \la{eq:orbitalmoment}
}
Here, the summation in $l$ does not include the two quasidegenerate bands.  There is finally a geometric contribution to the Zeeman interaction, which involves the two-by-two non-Abelian Berry connection  as well as the  zeroth-order band velocity: 
\e{\mathfrak{X}_{\beta,mn}{=}i\braket{u_m}{\partial u_n/\partial k_{\beta}},  \;\;v_{\alpha}:= \f{1}{\hbar}\p{\var}{k_{\alpha}},\la{berryconn}}
with $\{u_n\}_{n=1}^2$ the periodic component of Bloch functions in the quasidegenerate band. 
%which provides a geometric characterization of band wavefunctions along $\frako_0$.\\
%All matrix elements in \qq{eq:prop}{eq:orbitalmoment} are evaluated with wavefunctions of $\hat{H}_0$.

@@

\subsection{Recovering the Onsager-Lifshitz-Roth quantization rules in two limits}\la{sec:recoveronsager}

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics[nenciu] along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for quasidegenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \la{phaseindependentorbit}} 
In the second equality, we have applied \q{spinsplitwkb} and assumed that 
 $\delta \var$ is traceless (this assumption is  valid for splittings of a spin degeneracy\footnote{The Pauli spin-orbit coupling term ${\propto} \bs{\cdot} (\bE {\times} \bp)$ (that is derived from the non-relativistic limit of the Dirac equation) is traceless in any spin-symmetric (pseudo-)spinor basis.  Viewing the spin-orbit coupling as a small parameter in degenerate perturbation theory, the lowest-order splitting of any spin-degenerate energy level is always symmetric about the zeroth-order energy. This follows because the zeroth-order, spin-symmetric wavefunctions (associated to a spin-degenerate level) are tensor products of spin and spatial wavefunctions: $\ket{\Psi_{\pm}}{:}{=}\ket{{\pm}\hat{n}}{\otimes} \ket{\psi}$ with $\hat{n}$ a unit vector and  $\bs{\cdot} \hat{n} \ket{{\pm}\hat{n}}{=}{\pm}\ket{\pm \hat{n}}$. Hence, $\braopket{\Psi_{\pm}}{\bs{\cdot} (\bE {\times} \bp)}{\Psi_{\pm}}{=}{\pm}\hat{n}\cdot\braopket{\psi}{\bE {\times} \bp}{\psi}$. \textbf{Please generalize above proof to the case where H0 already includes centrosymmetric SOC. }}).  \q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule[onsager,lifshitz] for  two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth[rothII].\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S )$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits.[100page] } The perturbative inclusion of $B\calm_{\sma{+-}}$ results in  a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in  \s{sec:qo}.    \\

@@

For simplicity, we assume that the Pauli matrices in \q{eq:Rashba-Hamiltonian} correspond to spin: $s_i{=}\hbar \sigma_i/2$. Let us evaluate the effective Hamiltonian $\calh(\bk)$  along the cyclotron trajectory $\frako_0$ with radius $k$. In the  energy basis of the Rashba Hamiltonian $H_R$ [cf.\ \q{eq:Rashba-Hamiltonian}], 
\e{\calh(\bk)=-\alpha k\tau_z{+}\f{g_s}{2}\mu_{\sma{B}}B\tau_x {-} \f{eB\hbar}{2m} (\tau_x-I) \epsilon_{\alpha\beta} k_{\alpha}\nabla_{\beta}\theta,\la{rashbaeffham}}
where $\tau_i$ are a new set of Pauli matrices with $\tau_z{=}{+} 1$ and ${-}1$ corresponding to the outer and inner orbits, respectively. $e^{i\theta_{\bk}}{=}(ik_x{-}k_y)/k$ is the angle of the in-plane spin-splitting field $({-}\alpha k_y,\alpha k_x,0)$ of $H_R$, which winds once over $\frako_0$; the spin of our basis vectors likewise winds, in parallel (or anti-parallel) alignment with this field (see Fig. (\ref{fig:orbits})(b)). Terms proportional to the rate of spin rotation  ($\nabk \theta$) originate from the non-abelian Berry connection. The first term of  $\calh$ corresponds to the spin-orbit-induced splitting of the energy-degenerate bands;
the remaining terms may be interpreted as B-induced deformations of the bands.   It is well-known that the Zeeman term tends to polarize spin in the direction ($\vec{z}$) of $\bB$; remarkably, the non-abelian Berry term   also has a polarizing effect, but with opposite sign. These two effects completely cancel for free electrons ($m{=}m_0$, with $m_0$ being free electron mass.), but not generically in band systems. The orbital moment term $B M_z$, which generally couples the two-band subspace (of interest) to all other bands, is absent in any simplified two-band model.
%Observe that  that $\calh$ is actually time-independent, which originates from a  that is unbroken by the field (aligned with the rotational axis). Consequently,

@@

In the weak-hybridization regime: $l^2\delta S  \gg \pi|g_{\text{eff}}{-}2|$, we may expand \q{eq:Rashba-lambda} as a Laurent series in $B$:
\e{\lambda_{\pm}= \pm \f{l^2\delta S}{2} +\pi \pm \f{\pi^2}{4}\f{(g_{\text{eff}}-2)^2}{l^2\delta S} + O(B^2). \la{lambdasmallB}}
\q{eq:rule}, combined with the first two terms on the right-hand-side of \q{lambdasmallB}, may alternatively be derived from 
the standard Onsager-Lifshitz rules for two independent orbits, with a common $\pi$ Berry-phase correction for each orbit:\footnote{The intra-band Zeeman and orbital-moment couplings vanish due spacetime-inversion symmetry.[topoferm]} $l^2(S{\pm}\delta S)/2\pi {\in}\Z$. Higher-order terms in \q{lambdasmallB} reflect the field-induced hybridization of the spin-orbit-split orbits.

@@

Our quantization rule equally has utility where no closed-form analytic expression exists for the spectrum. Such is the case when we introduce a Dresselhaus spin-orbit coupling:
\begin{equation}
 H_{RD}(\bk)=\frac{{\hbar^2}k^2}{2m}+\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\beta(k_{x}\sigma_{x}-k_{y}\sigma_{y}).\la{hamRD}
\end{equation}
In Fig.\ \ref{fig:RD}(b-d), a comparison is made between the Landau levels obtained from exact numerical diagonalization of $H_{RD}(\bK)$ (by standard techniques reviewed in App. \ref{sec:numerical}) and those obtained from the quantization rule. The difference in energies tends to increase with smaller field [compare the left and right panels of   \fig{fig:RD}(b)] and shows a strong positive correlation with $l^2\delta S(\alpha,\beta)^2/S$ [compare \fig{fig:RD}(c) with \fig{fig:RD}(d)] for various choices of the spin-orbit parameters $(\alpha,\beta)$. This correlation may be anticipated from \q{eq:Rashba-exact}, which analytically demonstrates how terms involving $l^2\delta S^2/S$ modifies the quantization rule in the case where $\beta{=}0$. For our chosen parameters,  the difference in energies does not exceed $2\%$ of $\var_c$ as $l^2\delta S^2/S$ varies over $[0,0.75]$.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RD.png}
% alpha=0.05
% beta=0.02
% m=0.01
\caption{Rashba-Dresselhaus model. (a) Typical cyclotron orbits. (b) Comparison of Landau levels calculated by Eq. (\ref{eq:rule}) (black lines) and numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$ (red dots). Parameters for (b): $\alpha=7.5$($10^{5}$cm/s$\cdot\hbar$), $m=0.076m_e$, $\beta=3~(6)$($10^{5}$cm/s$\cdot\hbar$) for left (right) panels. (c) For various spin-orbit parameters $(\alpha,\beta)$ of the Rashba-Dresselhaus model and various field strength (satisfying $l^2S(E)\gg 1$), we plot  $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that lies closest to energy $E=0.4eV$, where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$. We also plot in (d) $l^2\delta S^2/S$, which has a strong positive relation with error in (c). Parameters for (c-d): $m=0.076m_e$. Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $\beta$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $l^2 S$($10^2$).\label{fig:RD}}
\end{figure}

\subsection{Rashba 2DEG with in-plane Zeeman field}\la{sec:inplanezeeman}

@@

\subsection{Relating the eigenvalue-degeneracy of $\A$ to the quasidegeneracy of Landau levels}\la{sec:relatedegeneracies}

>>>>

In the absence of magnetic breakdown, semiclassical quantization rules are well-known for Landau levels of  both nondegenerate [Onsager, Lifshitz] and degenerate [Roth, Mikitik, Topoferm, 100page] energy bands. Our work generalizes these rules to nearly-degenerate bands, whose orbits are split in $\bk$-space with a differential area  $\delta S$ that is much less than the average area $S$ of the orbits; we shall refer to $\delta S/S{\ll}1$ as the \textit{quasidegeneracy} condition. Quasidegenerate orbits commonly originate from spin-orbit-splitting of energy bands in noncentrosymmetric solids. We provide an illustrative model of a two dimensional electron gas (2DEG) with Rashba spin-orbit coupling:
\begin{equation}
H_R=\frac{{\hbar^2} k^2}{2m}+\alpha  (k_{x}\sigma_{y}-k_{y}\sigma_{x}).\label{eq:Rashba-Hamiltonian}
\end{equation}
At energies much greater than $m\alpha^2$, the differential area of the two spin-orbit-split orbits [see Fig. \ref{fig:orbits}(a)] is much smaller than their average area:\footnote{The expressions for $\delta S$ and $S$ in \q{deltaSvsS} are asymptotically valid for $E{\gg}m\alpha^2$.}
\e{\delta S = 2\pi k_E \f{2\alpha m}{\hbar^2} \ll S =\pi k_E^2; \;\; k_{E}:=\f{\sqrt{2m E}}{\hbar} .\label{deltaSvsS}}
$S$ is equivalently the area of the circular, constant-energy contour in the absence of spin-orbit coupling ($\alpha{=}0$).

@@

The quantization rule for Landau levels reflects the single-valuedness of the semiclassical (WKB) wavefunction over $\frako_0$,[Berry, Mount] i.e.,  the phase accumulated by an electron (over one cyclotron period $T_c$) is an integer multiple of $2\pi$: $(l^2S(E){+}\gamma)/2\pi{\in}\Z$. The leading-order phase   involves the classical action $S{:}{=}-\int_{0}^{T_c} k_x \dot{k}_y dt$  corresponding to $(K_x,K_y)$;[mount] $S$ has a dual interpretation as the signed area (in $\bk$-space) enclosed by $\frako_0$, as illustrated in Fig. \ref{fig:orbits}(b). $\gamma$ is a subleading Maslov correction [caustic] that generally equals $\pi$ times the rotation number $r$ of $\frako_0$ (i.e., the number of rotations made by the tangent vector over one cycle);[100page] for orbits that are deformable to a circle,  $r{=}{\pm} 1$. In the zeroth-order  solution, all Landau levels are spin-degenerate with a level spacing 
\e{ \var_c:=\f{2\pi}{l^2|\partial S/\partial E|}:=\f{\hbar^2}{ml^2}:=\f{h}{T_c}. \label{def:cyclotron}} 
Here, $\var_c$ and $T_c$ are respectively the cyclotron energy and period of the zeroth-order orbit, and $m$ is the effective mass. Henceforth, `degeneracy of Landau levels' should be understood as a spin degeneracy.

@@

To the first order, $\delta \hat{H}$ and the Zeeman interaction induce a subleading phase offset ($\lambda_a$) which generically differs for the two orthogonal spinor states (labelled $a{=}1,2$):
\begin{equation}
l^2S(E)+\lambda_a(E,B)+\gamma=2\pi n,\as  n\in \mathbb{Z},\;a=1,2. \label{eq:rule}
\end{equation}
This implies that the spin-degeneracy of Landau levels is split   by approximately  $\var_c|\lambda_1{-}\lambda_2|/2\pi$. The phase offsets $\lambda_a$ can be viewed as Floquet quasienergies of a time-periodic, perturbative Hamiltonian ($\calh$) that governs the dynamics of the spinor wavefunction.  $\lambda_a$ is obtained  from  eigenvalues $\{e^{i\lambda_a}\}_{a=1}^2$ of a time-evolution propagator -- the time-ordered exponential 
\begin{equation}
\A{:}{=}\overline{\exp}\left[-\frac{i}{\hbar}\int_0^{T_c} \calh(\bk(t)) dt\right],
\label{eq:prop}
\end{equation} 
with $\bk(t){\in}\frako_0$ a periodic function determined by the semiclassical equation of motion.   $\calh$ has the two-by-two matrix form:
\begin{equation}
  \calh=\delta \var +B\bigg(M_z -g_s\mu_{B}\f{s_z}{\hbar} + e\epsilon_{\alpha\beta}\mathfrak{X}_{\beta}v_{\alpha}\bigg):=\delta \var+B\calm,\label{eq:H1}
\end{equation}
$\delta \var$ is the projection of $\delta \hat{H}$ to the degenerate band subspace of $\hat{H}_0$. In the basis of energy-split bands (which we label by ${+}$ and $-$ for bands corresponding to the outer and inner orbit respectively), $\delta \var$ is a diagonal matrix whose diagonal entries ($\delta \var_+,\delta \var_-$, corresponding to outer and inner orbit respectively; $\delta \var_+<\delta \var_-$) are related to the hybridization parameter:\footnote{The neglected term on the right-hand side of  \q{spinsplitwkb} is smaller (than the kept term) by a factor $\delta v_{\pm}/v$, which is the relative change in band speed $v=|\boldsymbol{v}|=\hbar^{-1}|\partial \var/\partial \bk|$ induced by the spin-orbit coupling. This factor is assumed to be small.  For example in the Rashba model, $\delta v_{\pm}/v=m\alpha/\hbar k_E{\sim}\delta S/S{\ll}1$.}   
\e{-\int_0^{T_c} (\delta \var_+{-}\delta \var_-) \frac{dt}{\hbar} \approx l^2\delta S. \label{spinsplitwkb}}
%In the  Rashba model,   $\delta \var_{\pm}(E){=}{\pm} (\alpha/\hbar)\sqrt{2mE}$ is independent of $\bk$ due to a continuous rotational symmetry of $H_R$. \\

@@

The remaining terms in \q{eq:H1} describe a generalized Zeeman interaction $B\calm$ which is the first-order-in-$B$ correction to the Peierls-Onsager Hamiltonian $\var(\bK)$.[kohn,blount,roth] Included in this interaction is the Zeeman coupling to the spin ($g_s\mu_Bs_z/\hbar$) and orbital ($M_z$) magnetic moment of the quasidegenerate band; the latter is expressible in terms of matrix elements of the velocity operator $\boldsymbol{\Pi}_{mn}=i\langle\psi_m|[\hat{H}, \hat{\boldsymbol{r}}]|\psi_n\rangle/\hbar$ between Bloch wave functions $\psi_{n\bk}$:
\e{
M_z=\f{ie\hbar}{2}\sum_l'\f{\Pi_{x,ml}\Pi_{y,ln}-\Pi_{y,ml}\Pi_{x,ln}}{\epsilon_m-\epsilon_l}. \label{eq:orbitalmoment}
}
Here, the summation in $l$ does not include the two quasidegenerate bands.  There is finally a geometric contribution to the Zeeman interaction, which involves the two-by-two non-Abelian Berry connection  as well as the  zeroth-order band velocity: 
\e{\mathfrak{X}_{\beta,mn}{=}i\braket{u_m}{\partial u_n/\partial k_{\beta}},  \;\;v_{\alpha}:= \f{1}{\hbar}\p{\var}{k_{\alpha}},\label{berryconn}}
with $\{u_n\}_{n=1}^2$ the periodic component of Bloch functions in the quasidegenerate band. 
%which provides a geometric characterization of band wavefunctions along $\frako_0$.\\
%All matrix elements in \qq{eq:prop}{eq:orbitalmoment} are evaluated with wavefunctions of $\hat{H}_0$.

@@

\subsection{Recovering the Onsager-Lifshitz-Roth quantization rules in two limits}\label{sec:recoveronsager}

@@

\noi{i} If $|\var_+{-}\var_-|$ is everywhere (along $\frako_0$) much greater than $B|\calm_{+-}|$, then an electron would undergo adiabatic (band-conserving) dynamics[nenciu] along two independent, nondegenerate orbits. Such an adiabatic regime always exists as $B{\rightarrow} 0$ for quasidegenerate orbits that do not intersect, e.g, at a type-II Dirac point, which is the subject of the case studies in \s{sec:inplanezeeman} and \s{sec:rotsymmbreakdown}.
By  neglecting  $B\calm_{+-}$ in favor of intraband elements,  $\cala$ reduces to a diagonal matrix with  diagonal entries ($e^{i\lambda_{\pm}}$) given by:
\e{\lambda_{\sma{\pm}} = -\int_0^{T_c}\calh_{\sma{\pm\pm}}\f{dt}{\hbar}=\pm \f{1}{2}l^2\delta S - B\int_0^{T_c}\calm_{\sma{\pm\pm}}\f{dt}{\hbar}.  \label{phaseindependentorbit}} 
In the second equality, we have applied \q{spinsplitwkb} and assumed that 
 $\delta \var$ is traceless (this assumption is  valid for splittings of a spin degeneracy\footnote{The Pauli spin-orbit coupling term ${\propto} \bs{\cdot} (\bE {\times} \bp)$ (that is derived from the non-relativistic limit of the Dirac equation) is traceless in any spin-symmetric (pseudo-)spinor basis.  Viewing the spin-orbit coupling as a small parameter in degenerate perturbation theory, the lowest-order splitting of any spin-degenerate energy level is always symmetric about the zeroth-order energy. This follows because the zeroth-order, spin-symmetric wavefunctions (associated to a spin-degenerate level) are tensor products of spin and spatial wavefunctions: $\ket{\Psi_{\pm}}{:}{=}\ket{{\pm}\hat{n}}{\otimes} \ket{\psi}$ with $\hat{n}$ a unit vector and  $\bs{\cdot} \hat{n} \ket{{\pm}\hat{n}}{=}{\pm}\ket{\pm \hat{n}}$. Hence, $\braopket{\Psi_{\pm}}{\bs{\cdot} (\bE {\times} \bp)}{\Psi_{\pm}}{=}{\pm}\hat{n}\cdot\braopket{\psi}{\bE {\times} \bp}{\psi}$. \textbf{Please generalize above proof to the case where H0 already includes centrosymmetric SOC. }}).  \q{eq:rule}, combined with \q{phaseindependentorbit}, may be identified as the Onsager-Lifshitz rule[onsager,lifshitz] for  two independent, nondegenerate orbits of areas $S{\pm}\delta S$, and with subleading phase corrections that were first identified by Roth[rothII].\footnote{We remark that \q{phaseindependentorbit} neglects the contribution by band $-$ to the orbital moment of band $+$: $\delta M_z=il^{-2}\epsilon_{\ab}(\delta \var_+-\delta \var_-)\mathfrak{X}^{\alpha}_{+-}\mathfrak{X}^{\beta}_{-+}$. $\mathfrak{X}^{\alpha}_{+-}$, being an off-diagonal matrix element of the position operator, is generically of order a lattice constant $a$. We then estimate $\int_0^{T_c} B\delta M_z dt/\hbar=O(a^2 \delta S )$, which is negligible compared to $l^2\delta S/2$. Note also that the single-band orbital moment simply vanishes in some symmetry classes of orbits.[100page] } The perturbative inclusion of $B\calm_{\sma{+-}}$ results in  a correction to $\lambda_{\pm}$ [cf.\ \q{phaseindependentorbit}] that is expressible as a power series in $B$, with a leading term proportional to $B$; this will be confirmed in a subsequent case study [cf.\ \q{lambdasmallB}] and applied to understand beatings in quantum-oscillatory phenomena in  \s{sec:qo}.    \\

@@

For simplicity, we assume that the Pauli matrices in \q{eq:Rashba-Hamiltonian} correspond to spin: $s_i{=}\hbar \sigma_i/2$. Let us evaluate the effective Hamiltonian $\calh(\bk)$  along the cyclotron trajectory $\frako_0$ with radius $k$. In the  energy basis of the Rashba Hamiltonian $H_R$ [cf.\ \q{eq:Rashba-Hamiltonian}], 
\e{\calh(\bk)=-\alpha k\tau_z{+}\f{g_s}{2}\mu_{\sma{B}}B\tau_x {-} \f{eB\hbar}{2m} (\tau_x-I) \epsilon_{\alpha\beta} k_{\alpha}\nabla_{\beta}\theta,\label{rashbaeffham}}
where $\tau_i$ are a new set of Pauli matrices with $\tau_z{=}{+} 1$ and ${-}1$ corresponding to the outer and inner orbits, respectively. $e^{i\theta_{\bk}}{=}(ik_x{-}k_y)/k$ is the angle of the in-plane spin-splitting field $({-}\alpha k_y,\alpha k_x,0)$ of $H_R$, which winds once over $\frako_0$; the spin of our basis vectors likewise winds, in parallel (or anti-parallel) alignment with this field (see Fig. (\ref{fig:orbits})(b)). Terms proportional to the rate of spin rotation  ($\nabk \theta$) originate from the non-abelian Berry connection. The first term of  $\calh$ corresponds to the spin-orbit-induced splitting of the energy-degenerate bands;
the remaining terms may be interpreted as B-induced deformations of the bands.   It is well-known that the Zeeman term tends to polarize spin in the direction ($\vec{z}$) of $\bB$; remarkably, the non-abelian Berry term   also has a polarizing effect, but with opposite sign. These two effects completely cancel for free electrons ($m{=}m_0$, with $m_0$ being free electron mass.), but not generically in band systems. The orbital moment term $B M_z$, which generally couples the two-band subspace (of interest) to all other bands, is absent in any simplified two-band model.
%Observe that  that $\calh$ is actually time-independent, which originates from a  that is unbroken by the field (aligned with the rotational axis). Consequently,

@@

In the weak-hybridization regime: $l^2\delta S  \gg \pi|g_{\text{eff}}{-}2|$, we may expand \q{eq:Rashba-lambda} as a Laurent series in $B$:
\e{\lambda_{\pm}= \pm \f{l^2\delta S}{2} +\pi \pm \f{\pi^2}{4}\f{(g_{\text{eff}}-2)^2}{l^2\delta S} + O(B^2). \label{lambdasmallB}}
\q{eq:rule}, combined with the first two terms on the right-hand-side of \q{lambdasmallB}, may alternatively be derived from 
the standard Onsager-Lifshitz rules for two independent orbits, with a common $\pi$ Berry-phase correction for each orbit:\footnote{The intra-band Zeeman and orbital-moment couplings vanish due spacetime-inversion symmetry.[topoferm]} $l^2(S{\pm}\delta S)/2\pi {\in}\Z$. Higher-order terms in \q{lambdasmallB} reflect the field-induced hybridization of the spin-orbit-split orbits.

@@

Our quantization rule equally has utility where no closed-form analytic expression exists for the spectrum. Such is the case when we introduce a Dresselhaus spin-orbit coupling:
\begin{equation}
 H_{RD}(\bk)=\frac{{\hbar^2}k^2}{2m}+\alpha(k_{x}\sigma_{y}-k_{y}\sigma_{x})+\beta(k_{x}\sigma_{x}-k_{y}\sigma_{y}).\label{hamRD}
\end{equation}
In Fig.\ \ref{fig:RD}(b-d), a comparison is made between the Landau levels obtained from exact numerical diagonalization of $H_{RD}(\bK)$ (by standard techniques reviewed in App. \ref{sec:numerical}) and those obtained from the quantization rule. The difference in energies tends to increase with smaller field [compare the left and right panels of   \fig{fig:RD}(b)] and shows a strong positive correlation with $l^2\delta S(\alpha,\beta)^2/S$ [compare \fig{fig:RD}(c) with \fig{fig:RD}(d)] for various choices of the spin-orbit parameters $(\alpha,\beta)$. This correlation may be anticipated from \q{eq:Rashba-exact}, which analytically demonstrates how terms involving $l^2\delta S^2/S$ modifies the quantization rule in the case where $\beta{=}0$. For our chosen parameters,  the difference in energies does not exceed $2\%$ of $\var_c$ as $l^2\delta S^2/S$ varies over $[0,0.75]$.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RD.png}
% alpha=0.05
% beta=0.02
% m=0.01
\caption{Rashba-Dresselhaus model. (a) Typical cyclotron orbits. (b) Comparison of Landau levels calculated by Eq. (\ref{eq:rule}) (black lines) and numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$ (red dots). Parameters for (b): $\alpha=7.5$($10^{5}$cm/s$\cdot\hbar$), $m=0.076m_0$, $\beta=3~(6)$($10^{5}$cm/s$\cdot\hbar$) for left (right) panels. (c) For various spin-orbit parameters $(\alpha,\beta)$ of the Rashba-Dresselhaus model and various field strength (satisfying $l^2S(E)\gg 1$), we plot  $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that lies closest to energy $E=0.4eV$, where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian  $H_{RD}(\bK)$. We also plot in (d) $l^2\delta S^2/S$, which has a strong positive relation with error in (c). Parameters for (c-d): $m=0.076m_0$. Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $\beta$($1.5\times 10^{5}$cm/s$\cdot\hbar$), $l^2 S$($10^2$).\label{fig:RD}}
\end{figure}

\subsection{Rashba 2DEG with in-plane Zeeman field}\label{sec:inplanezeeman}

@@

\subsection{Relating the eigenvalue-degeneracy of $\A$ to the quasidegeneracy of Landau levels}\label{sec:relatedegeneracies}
